---
title: "ch2_estimation"
---

# 1.
The dataset `teengamb` concerns a study of teenage gambling in Britain. Fit a regression model with the expenditure on gambling as the response and the sex, status, income, and verbal score as predictors. 

```{r 1}
library(faraway)
library(tidyverse)

data(teengamb)
teengamb$sex <- factor(teengamb$sex)
levels(teengamb$sex) <- c("male", "female")

tgamb_lm <- lm(gamble ~ sex + status + income + verbal,
           data = teengamb)
sumary(tgamb_lm)
```

## a.
What percentage of variation in the response is explained by these predictors?

R-squared explains to what extent the variance of one variable explains the variance of the second variable. In this model, 53% of the observed variation can be explained by those predictors. 

## b.
Which observation has the largest (positive) residual? Give the case number. 

```{r 1b}
which.max(resid(tgamb_lm))
```

A residual is the vertical distance between a point and the regression line. The observation with the largest postive residual is case number 24.

## c.
Compute the mean and median of the residuals. 

```{r 1c}
summary(resid(tgamb_lm))
```

The mean is  0 and the median is -1.451. 

## d. 
Compute the correlation of the residuals with the fitted values. 

```{r 1d}
cor(fitted(tgamb_lm), resid(tgamb_lm))
```

The correlation is quite small, at -1.071e-16. 

## e. 
Compute the correlation of the residuals with the income. 

```{r 1e}
cor(teengamb$income, resid(tgamb_lm))
```

The correlation is even smaller, at -7.242e-17.




