---
title: "ch2_estimation"
---

# 1.
The dataset `teengamb` concerns a study of teenage gambling in Britain. Fit a regression model with the expenditure on gambling as the response and the sex, status, income, and verbal score as predictors. 

```{r 1}
library(faraway)
library(magrittr)
library(tidyverse)

data(teengamb)
teengamb$sex <- factor(teengamb$sex)
levels(teengamb$sex) <- c("male", "female")

tgamb_lm <- lm(gamble ~ sex + status + income + verbal,
           data = teengamb)
summary(tgamb_lm)
```

## a.
What percentage of variation in the response is explained by these predictors?

R-squared explains to what extent the variance of one variable explains the variance of the second variable. In this model, 53% of the observed variation can be explained by those predictors. 

## b.
Which observation has the largest (positive) residual? Give the case number. 

```{r 1b}
which.max(resid(tgamb_lm))
```

A residual is the vertical distance between a point and the regression line. The observation with the largest postive residual is case number 24.

## c.
Compute the mean and median of the residuals. 

```{r 1c}
summary(resid(tgamb_lm))
```

The mean is  0 and the median is -1.451. 

## d. 
Compute the correlation of the residuals with the fitted values. 

```{r 1d}
cor(fitted(tgamb_lm), resid(tgamb_lm))
```

The correlation is quite small, at -1.071e-16. 

## e. 
Compute the correlation of the residuals with the income. 

```{r 1e}
cor(teengamb$income, resid(tgamb_lm))
```

The correlation is even smaller, at -7.242e-17.

## f. 
For all other predictors held constant, what would be the difference in predicted expenditure on gambling for a male compared to a female?

Looking back at `summary(tgamb_lm)`, the coefficient of `sexfemale` is -22.12. That means, the predicted expenditure drops $22.12 for women when compared to men. 

# 2.
The dataset `uswages` is drawn from a sample of the Current Population Survey in 1988. Fit a model with weekly wages as the response and years of education and experience as predictors. Report and give a simple interpretation to the regression coefficient for years of education. Now fit the same model but with logged weekly wages. Give an interpretation to the regression coefficient for years of education. Which interpretation is more natural?

```{r 2}
data(uswages)

# make negative experience values 0
uswages$exper %<>% ifelse(. < 0, 0, .)

wage_lm <- lm(wage ~ educ + exper, data = uswages)
sumary(wage_lm)

log_wage_lm <- lm(log(wage) ~ educ + exper, data = uswages)
sumary(log_wage_lm)
```

For every year of education, weekly wages should increase by $51.12 while considering experience. When the response variable is logged, for every year of education, weekly wages should increase by 9% considering experience. 

